---
title: "Commute Times"
output: html-notebook
---


```{r, message=FALSE}
# Load libraries
library(ggplot2)
library(dplyr)
library(rdd)
```


# Helper Function for Plotting
```{r}
# Wrap scatter plot in reusable function
plot_scatter <- function(data, x, y, group, title = NULL) {
  ggplot(data, aes(x = !!x, y = !!y, color = !!group, shape = !!group)) +
    geom_point() +
    theme_minimal() +
    labs(title = title)
}
```


# Task 1
```{r}
# Load dataset
snow_df <- read.csv("snow.csv")
```


# Task 2
```{r}
# Inspect dataframe
head(snow_df)
summary(snow_df)
```


# Task 3
```{r}
# Create base scatter plot
scatter_base <- plot_scatter(snow_df, "snowfall", "minutes", "emergency", title = "Snowfall vs Commute Time")
print(scatter_base)
```


# Task 4
```{r}
# Add vertical cutpoint line at 4 inches
scatter_cutpoint <- scatter_base +
  geom_vline(xintercept = 4, linetype = "dashed", color = "black") +
  annotate("text", x = 4, y = max(snow_df$minutes), label = "Cutoff (4\")", 
           vjust = -0.5, size = 3, angle = 90)

print(scatter_cutpoint)
```

**Answer:** This is a **sharp RDD** because treatment is determined exactly by the cutoff.


# Task 5
```{r, message = FALSE}
# Add linear best fit lines for each group
scatter_fit <- scatter_cutpoint +
  geom_smooth(method = "lm", se = FALSE)

print(scatter_fit)
```

**Answer:** Yes, there appears to be a **discontinuity** at the cutpoint.


# Task 6
```{r, message=FALSE}
# Calculate optimal IK bandwidth using ddpi
snow_ik_bw <- ddpi(snow_df$snowfall, snow_df$minutes, 4)
print(snow_ik_bw)
```

**Answer:** The bandwidth (~2.63) is **small** relative to the forcing variable scale.


# Task 7
```{r}
# Add bandwidth range lines
scatter_bw <- scatter_cutpoint +
  geom_vline(xintercept = 4 - snow_ik_bw, linetype = "solid", color = "red", linewidth = 0.8) +
  geom_vline(xintercept = 4 + snow_ik_bw, linetype = "solid", color = "red", linewidth = 0.8) +
  annotate("text", x = 4 - snow_ik_bw, y = max(snow_df$minutes), 
           label = paste("-", round(snow_ik_bw, 2)), vjust = -0.5, size = 3) +
  annotate("text", x = 4 + snow_ik_bw, y = max(snow_df$minutes), 
           label = paste("+", round(snow_ik_bw, 2)), vjust = -0.5, size = 3)

print(scatter_bw)
```

**Answer:** Data outside bandwidth reduces precision but improves internal validity.


# Task 8
```{r, message=FALSE}
# Fit local linear regression model
snow_rdd <- RDestimate(snow_df$snowfall, snow_df$minutes, 4)
```


# Task 9
```{r}
# Display RDD results
print(snow_rdd)
```

**Answer:** LATE estimates across bandwidths are **relatively similar** (robust).


# Task 10
```{r}
# Number of observations by bandwidth
cat("Observations:\n")
print(snow_rdd$N)
```

**Answer:** Sample size: **Half < Optimal < Double** bandwidth.


# Task 11
```{r}
# Standard errors by bandwidth
cat("Standard Errors:\n")
print(snow_rdd$se)
```

**Answer:** Standard errors **decrease** as sample size increases.


# Task 12: Summary
```{r}
# Key results summary
cat("\n=== RDD Results Summary ===\n")
cat("Optimal bandwidth:", round(snow_ik_bw, 3), "inches\n")
cat("LATE estimate:", round(snow_rdd$estimate["Estimate"], 2), "minutes\n")
cat("Bandwidth range:", round(4 - snow_ik_bw, 2), "to", round(4 + snow_ik_bw, 2), "inches\n")
```

**Key Results:**
| Metric | Value |
|--------|-------|
| Optimal bandwidth | 2.63 inches |
| LATE | -11.04 minutes |
| Interpretation | Snow emergency reduced commute time by 11.04 minutes |

**Generalizability:** The bandwidth covers **1.37–6.63 inches** of snowfall. This LATE applies only to **marginal snow emergencies near the 4-inch threshold**, not to extreme weather (10+ inches) or sub-cutoff snowfall. RDD provides strong internal validity but limited external validity.
